The inefficacy of superficial similarities for improving instructor-student relationships

### 2021-04-27 outputs/paper/paper.Rmd
actual last edits.

```{r, warning=FALSE, message=FALSE, echo=FALSE}
expdata <- isrdata %>%
  filter(t1_consent == 1 # Only include if instructor consented to first survey
         & t1_complete == 1 # Only included if instructor completed first survey
         & t1_timer_consent_3 > 1 # Only include if instructor read consent
         & s1_consent == 1 # Only include if student consented to first survey
         & s1_timer_consent_3 > 1 # Only include if student read consent
         & s1_fullsurvey == 1 # Only include if the student completed the whole first survey
         & t2_complete == 1 # Only include if the instructor completed the whole second survey
         # The rest of the lines make sure that the survey taker looked at each page for more than three seconds
         & t1_timer_gtky1_3 > 10
         & t1_timer_gtky2_3 > 10
         & t1_timer_gtky3_3 > 10
         & t1_timer_demos1_3 > 10
         & t1_timer_demos2_3 > 10
         & s1_timer_gtky1_3 > 10
         & s1_timer_gtky2_3 > 10
         & s1_timer_gtky3_3 > 10
         & s1_timerpre1_3 > 10
         & s1_timerpre2_3 > 10
         & (s1_timercontrolfeedback_3 > 10 | is.na(s1_timercontrolfeedback_3))
         & (s1_timercontrolresponse_3 > 10 | is.na(s1_timercontrolresponse_3))
         & (s1_timertreatmentfeedba_3 > 10 | is.na(s1_timertreatmentfeedba_3))
         & (s1_timertreatmentrespon_3 > 10 | is.na(s1_timertreatmentrespon_3))
         & s1_timerpost1_3 > 10
         & s1_timerpost2_3 > 10
         & s1_timerdemo1_3 > 10
         & (s2_timer1a_3 > 10 | is.na(s2_timer1a_3))
         & (s2_timer1b_3 > 10 | is.na(s2_timer1b_3))
         & (s2_timer2a_3 > 10 | is.na(s2_timer2a_3))
         & (s2_timer3a_3 > 10 | is.na(s2_timer3a_3))
         & (s2_timer3b_3 > 10 | is.na(s2_timer3b_3))
         & (s2_timer3c_3 > 10 | is.na(s2_timer3c_3))
         & (s2_timer4_3 > 10 | is.na(s2_timer4_3))
         & t2_timer > 10
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         )

# Alternatively: expdata <- isrdata %>% filter(use == 1) %>% # Select only the cases the original study used.

@@

expdata <- expdata %>%
  mutate(t2_finalexamobj = ifelse(obj_exam == 1, t2_finalexam, NA))
```

@@

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
s1_sim_scale <- expdata %>%
  dplyr::select(s1_sim1, s1_sim2, s1_sim3, s1_sim4, s1_sim5, s1_sim6) %>%
  drop_na()
cronbach.alpha(s1_sim_scale)

@@

The data, in a .dta format, and Stata code for the @citeoriginal paper are available on the Inter-University Consortium for Political and Social Research [@Robinson_Scott_Gottfried_2020]. The dataset contains 36,838 observations of 653 variables.

@@

instructorrace <- expdata %>%
  dplyr::select(teacherid,
         t1_race_1, t1_race_2, t1_race_3, t1_race_4, t1_race_5, t1_race_6, t1_race_7, t_multi_race) %>%
  unique() %>%
  dplyr::summarize('White/Caucasian' = sum(t1_race_1, na.rm = TRUE)/n(),
            'Black or African American' = sum(t1_race_2, na.rm = TRUE)/n(),
            'Hispanic American or Latino/a' = sum(t1_race_3, na.rm = TRUE)/n(),
            'Asian or Pacific Islander' = sum(t1_race_4, na.rm = TRUE)/n(),
            'American Indian or Alaskan' = sum(t1_race_5, na.rm = TRUE)/n(),
            'Middle Eastern' = sum(t1_race_6, na.rm = TRUE)/n(),
            'Other Race' = sum(t1_race_7, na.rm = TRUE)/n(),
            'Multiple' = sum(t_multi_race > 1, na.rm = TRUE)/n())
studentrace <- expdata %>%
  dplyr::select(id,
         s1_race_1, s1_race_2, s1_race_3, s1_race_4, s1_race_5, s1_race_6, s1_race_7, s_multi_race) %>%
  unique() %>%
  dplyr::summarize('White/Caucasian' = sum(s1_race_1, na.rm = TRUE)/n(),
            'Black or African American' = sum(s1_race_2, na.rm = TRUE)/n(),
            'Hispanic American or Latino/a' = sum(s1_race_3, na.rm = TRUE)/n(),
            'Asian or Pacific Islander' = sum(s1_race_4, na.rm = TRUE)/n(),
            'American Indian or Alaskan' = sum(s1_race_5, na.rm = TRUE)/n(),
            'Middle Eastern' = sum(s1_race_6, na.rm = TRUE)/n(),
            'Other Race' = sum(s1_race_7, na.rm = TRUE)/n(),
            'Multiple' = sum(s_multi_race > 1, na.rm = TRUE)/n())

bind_cols(as_tibble(cbind(nms = names(instructorrace), t(instructorrace))),
          as_tibble(cbind(nms = names(studentrace), t(studentrace))))%>%
  dplyr::select(-nms...3) %>%
  pivot_longer(!nms...1, names_to = "group", values_to = "portion") %>%
  mutate(nms...1 = fct_relevel(nms...1, 
                               c('Hispanic American or Latino/a',
                                 'White/Caucasian',
                                 'Asian or Pacific Islander',
                                 'Middle Eastern',
                                 'Black or African American', 
                                 'American Indian or Alaskan',
                                 'Other Race',
                                 'Multiple')),
         group = fct_relevel(group, c("V2...4", "V2...2"))) %>%
  ggplot() +
  geom_col(aes(x = nms...1, y = as.numeric(portion), fill = group), position = "dodge") +
  labs(x = "Self-identified race/ethnicity", y = "Proportion of sample", 
       fill = "Group", title = "Self-identified racial proportions differ for instructors and students") +
  scale_fill_manual(values = c("#c8b5f7", "#f9b5c9"), labels = c("Students", "Instructors")) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 11)) +
  theme_minimal()
```

@@

```{r classsizegraph, echo=FALSE, warning=FALSE, message=FALSE, fig.dim=c(8.5,3), fig.cap="Histogram illustrating class sizes", out.width="100%", fig.align="center"}
n_instructor <- expdata %>%
  dplyr::select(teacherid, n_course) %>%
  unique()

n_by_instructor <- n_instructor %>%
  ggplot(aes(x = n_course)) +
  geom_histogram(binwidth = 5, fill = "#e9fcbe", color = "#f48cc6") +
  geom_vline(xintercept = mean(n_instructor %>% pull(n_course), na.rm = TRUE),
             col = "#868cf7",
             lwd = .5) +
  labs(x = "Class size", y = "Number of classes") +
  theme_minimal()

n_student <- expdata %>%
  dplyr::select(id, n_course) %>%
  unique()

n_by_student <- n_student %>%
  ggplot(aes(x = n_course)) +
  geom_histogram(binwidth = 5, fill = "#e9fcbe", color = "#f48cc6") +
  geom_vline(xintercept = mean(n_student %>% pull(n_course), na.rm = TRUE),
             col = "#868cf7",
             lwd = .5) +
  labs(x = "Class size", y = "Number of students in study") +
  theme_minimal()

combined <- plot_grid(n_by_instructor, n_by_student,
          ncol = 2)

title <- ggdraw() + 
  draw_label(
    "Some classes are very large (class size means marked)",
    x = 0,
    hjust = 0
  ) +
  theme(
    plot.margin = margin(0, 0, 0, 7)
  )

plot_grid(title,
          combined,
          legend,
          ncol = 1,
          rel_heights = c(.09, 2))
```

@@

Interestingly, there is a large difference between how students rate perceived similarity and how instructors rate perceived similarity. Students, on average, rate end of term similarity at 3.6, while instructors give an average rating of 2.7. Students give an average end-of-term ISR rating of 4.1, while instructors give an average of only 3.5. This could be due to more positive student perceptions, or due to students giving higher ratings for equivalent perceptions. Both groups rate ISR more highly than similarity, but it is difficult to compare the weight of these scores as the scales are not necessarily measuring in the same units.

```{r continuousoutcomes, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6, fig.height= 10, fig.cap="Histograms illustrating continous and ordinal key outcomes"}
s1_simplot <- expdata %>% 
  ggplot(aes(x = s1_sim, fill = treatment)) +
  geom_histogram(position = "dodge") +
  labs(x = "Initial student similarity perception", y = "Frequency", fill = "Group") +
  scale_x_continuous(limits = c(.8,5.2)) +
  scale_fill_manual(values = c("#52f2b9", "#868cf7")) +
  theme_minimal()

s1_tsrplot <- expdata %>% 
  ggplot(aes(x = s1_tsr, fill = treatment)) +
  geom_histogram(position = "dodge") +
  labs(x = "Student ISR anticipation", y = "Frequency") +
  scale_x_continuous(limits = c(.8,5.2)) +
  scale_fill_manual(values = c("#52f2b9", "#868cf7")) +
  theme_minimal()

s2_simplot <- expdata %>% 
  ggplot(aes(x = s2_sim, fill = treatment)) +
  geom_histogram(position = "dodge") +
  labs(x = "End-of-term student similarity perception", y = "Frequency") +
  scale_x_continuous(limits = c(.8,5.2)) +
  scale_fill_manual(values = c("#52f2b9", "#868cf7")) +
  theme_minimal()

s2_tsrplot <- expdata %>% 
  ggplot(aes(x = s2_tsr, fill = treatment)) +
  geom_histogram(position = "dodge") +
  labs(x = "End-of-term student ISR perception", y = "Frequency") +
  scale_x_continuous(limits = c(.8,5.2)) +
  scale_fill_manual(values = c("#52f2b9", "#868cf7")) +
  theme_minimal()

t2_sim1plot <- expdata %>% 
  ggplot(aes(x = t2_sim1, fill = treatment)) +
  geom_histogram(position = "dodge") +
  labs(x = "End-of-term instructor similarity perception", y = "Frequency") +
  scale_x_continuous(limits = c(.8,5.2)) +
  scale_fill_manual(values = c("#52f2b9", "#868cf7")) +
  theme_minimal()

t2_tsrplot <- expdata %>% 
  ggplot(aes(x = t2_tsr, fill = treatment)) +
  geom_histogram(position = "dodge") +
  labs(x = "End-of-term instructor ISR perception", y = "Frequency") +
  scale_x_continuous(limits = c(.8,5.2)) +
  scale_fill_manual(values = c("#52f2b9", "#868cf7")) +
  theme_minimal()

gradeplot <- expdata %>% 
  ggplot(aes(x = grade, fill = treatment)) +
  geom_histogram(position = "dodge") +
  labs(x = "Course grade", y = "Frequency") +
  scale_x_continuous(limits = c(-.1,4.5)) +
  scale_fill_manual(values = c("#52f2b9", "#868cf7")) +
  theme_minimal()

std_gradeplot <- expdata %>% 
  ggplot(aes(x = std_grade, fill = treatment)) +
  geom_histogram(position = "dodge") +
  labs(x = "Standardized course grade", y = "Frequency") +
  scale_x_continuous(limits = c(-4.5,3)) +
  scale_fill_manual(values = c("#52f2b9", "#868cf7")) +
  theme_minimal()

t2_finalexamplot <- expdata %>% 
  ggplot(aes(x = t2_finalexam, fill = treatment)) +
  geom_histogram(position = "dodge") +
  labs(x = "Final grade", y = "Frequency") +
  scale_x_continuous(limits = c(-.1,4.5)) +
  scale_fill_manual(values = c("#52f2b9", "#868cf7")) +
  theme_minimal()

combined <- plot_grid(s1_simplot + theme(legend.position = "none"), s1_tsrplot + theme(legend.position = "none"),
          s2_simplot + theme(legend.position = "none"), s2_tsrplot + theme(legend.position = "none"),
          t2_sim1plot + theme(legend.position = "none"), t2_tsrplot + theme(legend.position = "none"),
          std_gradeplot + theme(legend.position = "none"), t2_finalexamplot + theme(legend.position = "none"),
          ncol = 2)

legend <- get_legend(s1_simplot + 
                       guides(color = guide_legend(nrow = 1)) +
                       theme(legend.position = "bottom"))

title <- ggdraw() + 
  draw_label(
    "Most key outcomes unaffected by treatment",
    x = 0,
    hjust = 0
  ) +
  theme(
    plot.margin = margin(0, 0, 0, 7)
  )

plot_grid(title,
          combined,
          legend,
          ncol = 1,
          rel_heights = c(.1, 2, .1))
```

@@

From the 36,838 observations, or potential units of study (corresponding to all undergraduate student records of the university), only 2,273 were used in analysis. Units of study were excluded because the instructor did not participate in the study (33,483); the instructor or course wasn't actually eligible (96); the instructor administered the wrong survey (31); the student did not consent  (116); the student did not complete the first survey (304); the instructor did not complete the second survey (107); or one participant did not spend at more than a second reading the consent page and more than ten seconds reading each page with five questions or more (428). The time limits are somewhat arbitrary, and different choices could have different analysis results.

@@

$X_{1i}$ is a vector of student-level covariates (student ISR anticipation, gender, CGPA, year of study). Different covariates are used to model different outcomes. The original study used gender and CGPA are used as controls for grade based outcomes only. @citeoriginal state this is because females earn higher grades generally. This paper uses them as controls for the majority of outcomes. ISR anticipation is used as a control only for end-of-term student ISR perception.

$X_{2j}$ is a vector of instructor-level covariates (class size). Different covariates are used to model different outcomes. The original study consider instructor fixed effects instead. Because each participating instructor taught one class in the study, the class sizes correspond directly to the instructor IDs, so it was inadvisable to use both. This paper selected class size because of the larger potential for transferable insights.

$\epsilon_{ij}$ is a standard error. As this paper does not use clustered standard errors like @citeoriginal, there are slight differences in standard error and p-values when compared with the original paper. This does not affect the regression coefficients.

@@

Equation \@ref(eq:linmodel) is @citeoriginal's linear model, used for continuous outcomes. This include complete scale outcomes: initial student similarity perception, end-of-term student similarity perception, end-of-term student ISR perception, and end-of-term instructor ISR perception. Because these scales are created using 6-7 questions answered on a scale of 1 to 5, they have a total of 25 or 30 possible values, they can be treated as continuous variables and modeled using linear functions. It also includes grade-based outcomes: course grade and objectively graded exam grade, both of which are shown on a 4.0 GPA scale.

@@

Equation \@ref(eq:ordlogmodel) is @citeoriginal's ordinal logistic model, used for the ordinal outcome, which is end-of-term instructor similarity perception. Because the instructors were only asked to answer one five-point scale question about similarity, this outcome cannot be treated the same way as the full-scale outcomes.

@@

When creating additional models, I focused on three key outcomes: standardized course grade, end-of-term student perceived ISR scale, and end-of-term student perceived similarity scale. I selected these because interventions like @citeoriginal's are aimed at improving student outcomes (like course performance), and it is much more likely to detect an impact in the immediate course outcomes than outcomes from subsequent terms; and because previous studies show more detectable impact on student perception than instructor perception. Equation \@ref(eq:mylinmodel) shows the linear model used. $X_{3jk}$ represents student-instructor commonality covariates (matching racial self-ID, matching gender self-ID, age difference).

@@

In Table \@ref(tab:modeltable), Model 1 shows that the treatment had a statistically significant effect on initial student similarity perception ($\beta_{1}$ = 0.194, 95% CI [0.136, 0.253]). Similarly, Model 2 that the treatment had a statistically significant effect on  student end-of-term similarity perception ($\beta_{1}$ = 0.11, 95% CI [0.038, 0.183]), though this effect size is smaller. Both coefficients are minimal on a scale of 1 to 5. The adjusted R-squared values, which indicate the proportion of variation in the outcome that can be explained by the treatment and covariates, are low (0.032 and 0.022). These models are helpful for analyzing treatment effect, but they do not aim to be and are not useful predictors of student similarity perception, not even within the sample.

@@

Models 3 and 4 show that student ISR perception is significantly associated course grade, while instructor ISR perception is significantly associated with course grade and final grade. Instructor perception is more strongly associated ($\Gamma$ = 0.34, 95% CI [0.295, 0.385] for course grade; $\Gamma$ = 0.643, 95% CI [0.642, 0.759] for final grade) than student perception ($\Gamma$ = 0.115, 95% CI [0.064, 0.164] for course grade).

@@

```{r, echo=FALSE, warning=FALSE, message=FALSE}
expdata <- expdata %>%
  mutate(race_match = ifelse(t1_race_1 == s1_race_1
                             & t1_race_2 == s1_race_2 
                             & t1_race_3 == s1_race_3 
                             & t1_race_4 == s1_race_4 
                             & t1_race_5 == s1_race_5 
                             & t1_race_6 == s1_race_6 
                             & t1_race_7 == s1_race_7,
                             1,
                             0)) %>%
  mutate(gend_match = ifelse(t_female == s1_female,
                             1,
                             0)) %>%
  mutate(age_dif = t_age - s1_age)
```

>>>>

```{r, warning=FALSE, message=FALSE, echo=FALSE}
expdata <- isrdata %>%
  filter(t1_consent == 1 # Only include if instructor consented to first survey
         & t1_complete == 1 # Only included if instructor completed first survey
         & t1_timer_consent_3 > 1 # Only include if instructor read consent
         & s1_consent == 1 # Only include if student consented to first survey
         & s1_timer_consent_3 > 1 # Only include if student read consent
         & s1_fullsurvey == 1 # Only include if the student completed the whole first survey
         & t2_complete == 1 # Only include if the instructor completed the whole second survey
         # Make sure that the survey taker looked at each page for more than three seconds
         & t1_timer_gtky1_3 > 10
         & t1_timer_gtky2_3 > 10
         & t1_timer_gtky3_3 > 10
         & t1_timer_demos1_3 > 10
         & t1_timer_demos2_3 > 10
         & s1_timer_gtky1_3 > 10
         & s1_timer_gtky2_3 > 10
         & s1_timer_gtky3_3 > 10
         & s1_timerpre1_3 > 10
         & s1_timerpre2_3 > 10
         & (s1_timercontrolfeedback_3 > 10 | is.na(s1_timercontrolfeedback_3))
         & (s1_timercontrolresponse_3 > 10 | is.na(s1_timercontrolresponse_3))
         & (s1_timertreatmentfeedba_3 > 10 | is.na(s1_timertreatmentfeedba_3))
         & (s1_timertreatmentrespon_3 > 10 | is.na(s1_timertreatmentrespon_3))
         & s1_timerpost1_3 > 10
         & s1_timerpost2_3 > 10
         & s1_timerdemo1_3 > 10
         & (s2_timer1a_3 > 10 | is.na(s2_timer1a_3))
         & (s2_timer1b_3 > 10 | is.na(s2_timer1b_3))
         & (s2_timer2a_3 > 10 | is.na(s2_timer2a_3))
         & (s2_timer3a_3 > 10 | is.na(s2_timer3a_3))
         & (s2_timer3b_3 > 10 | is.na(s2_timer3b_3))
         & (s2_timer3c_3 > 10 | is.na(s2_timer3c_3))
         & (s2_timer4_3 > 10 | is.na(s2_timer4_3))
         & t2_timer > 10
         # Remove specific drop reasons
         & t_drop_reasons != "Teacher administered GTKY survey mistakenly at end of semester"
         & t_drop_reasons != "Taught online course"
         & t_drop_reasons != "Administered to only graduate students"
         )

@@

# This creates a variable that is the final grade if the final was an objectively graded exam
# and otherwise NA
expdata <- expdata %>%
  mutate(t2_finalexamobj = ifelse(obj_exam == 1, t2_finalexam, NA))
```

@@

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
# Conduct tests for scale reliability
s1_sim_scale <- expdata %>%
  dplyr::select(s1_sim1, s1_sim2, s1_sim3, s1_sim4, s1_sim5, s1_sim6) %>%
  drop_na()
cronbach.alpha(s1_sim_scale)

@@

The data, in a .dta format, and Stata code for the @citeoriginal paper are available on the Inter-University Consortium for Political and Social Research [@Robinson_Scott_Gottfried_2020]. The dataset contains 36,838 observations of 653 variables. While there are many additional variables, including a wider range of survey questions targeting interest level, enjoyment, and confidence, this paper focuses on the covariates and key outcomes that are discussed in the original paper.

@@

# Summarise instructor race/ethnicity data
instructorrace <- expdata %>%
  dplyr::select(teacherid,
         t1_race_1, t1_race_2, t1_race_3, t1_race_4, t1_race_5, t1_race_6, t1_race_7, t_multi_race) %>%
  unique() %>%
  dplyr::summarize('White/Caucasian' = sum(t1_race_1, na.rm = TRUE)/n(),
            'Black or African American' = sum(t1_race_2, na.rm = TRUE)/n(),
            'Hispanic American or Latino/a' = sum(t1_race_3, na.rm = TRUE)/n(),
            'Asian or Pacific Islander' = sum(t1_race_4, na.rm = TRUE)/n(),
            'American Indian or Alaskan' = sum(t1_race_5, na.rm = TRUE)/n(),
            'Middle Eastern' = sum(t1_race_6, na.rm = TRUE)/n(),
            'Other Race' = sum(t1_race_7, na.rm = TRUE)/n(),
            'Multiple' = sum(t_multi_race > 1, na.rm = TRUE)/n())

# summarise student race/ethnicity data
studentrace <- expdata %>%
  dplyr::select(id,
         s1_race_1, s1_race_2, s1_race_3, s1_race_4, s1_race_5, s1_race_6, s1_race_7, s_multi_race) %>%
  unique() %>%
  dplyr::summarize('White/Caucasian' = sum(s1_race_1, na.rm = TRUE)/n(),
            'Black or African American' = sum(s1_race_2, na.rm = TRUE)/n(),
            'Hispanic American or Latino/a' = sum(s1_race_3, na.rm = TRUE)/n(),
            'Asian or Pacific Islander' = sum(s1_race_4, na.rm = TRUE)/n(),
            'American Indian or Alaskan' = sum(s1_race_5, na.rm = TRUE)/n(),
            'Middle Eastern' = sum(s1_race_6, na.rm = TRUE)/n(),
            'Other Race' = sum(s1_race_7, na.rm = TRUE)/n(),
            'Multiple' = sum(s_multi_race > 1, na.rm = TRUE)/n())

# bind the two summaries and present them in a bar chart
bind_cols(as_tibble(cbind(nms = names(instructorrace), t(instructorrace))),
          as_tibble(cbind(nms = names(studentrace), t(studentrace))))%>%
  dplyr::select(-nms...3) %>%
  pivot_longer(!nms...1, names_to = "group", values_to = "portion") %>%
  mutate(nms...1 = fct_relevel(nms...1, 
                               c('Hispanic American or Latino/a',
                                 'White/Caucasian',
                                 'Asian or Pacific Islander',
                                 'Middle Eastern',
                                 'Black or African American', 
                                 'American Indian or Alaskan',
                                 'Other Race',
                                 'Multiple')),
         group = fct_relevel(group, c("V2...4", "V2...2"))) %>%
  ggplot() +
  geom_col(aes(x = nms...1, y = as.numeric(portion), fill = group), position = "dodge") +
  labs(x = "Self-identified race/ethnicity", y = "Proportion of sample", 
       fill = "Group", title = "Self-identified racial proportions differ for instructors and students") +
  scale_fill_manual(values = c("#c8b5f7", "#f9b5c9"), labels = c("Students", "Instructors")) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 11)) +
  theme_minimal()
```

@@

```{r classsizegraph, echo=FALSE, warning=FALSE, message=FALSE, fig.dim=c(8.5,3), fig.cap="Histogram illustrating class sizes", out.width="100%", fig.align="center"}

# Summarise course sizes from instructor POV
n_instructor <- expdata %>%
  dplyr::select(teacherid, n_course) %>%
  unique()

# Plot instructor POV course sizes
n_by_instructor <- n_instructor %>%
  ggplot(aes(x = n_course)) +
  geom_histogram(binwidth = 5, fill = "#e9fcbe", color = "#f48cc6") +
  geom_vline(xintercept = mean(n_instructor %>% pull(n_course), na.rm = TRUE),
             col = "#868cf7",
             lwd = .5) +
  labs(x = "Class size", y = "Number of classes") +
  theme_minimal()

# Summarise course sizes from a student POV
n_student <- expdata %>%
  dplyr::select(id, n_course) %>%
  unique()

# Plot student POV summary
n_by_student <- n_student %>%
  ggplot(aes(x = n_course)) +
  geom_histogram(binwidth = 5, fill = "#e9fcbe", color = "#f48cc6") +
  geom_vline(xintercept = mean(n_student %>% pull(n_course), na.rm = TRUE),
             col = "#868cf7",
             lwd = .5) +
  labs(x = "Class size", y = "Number of students in study") +
  theme_minimal()

# Combine the plots in a grid
combined <- plot_grid(n_by_instructor, n_by_student,
          ncol = 2)
title <- ggdraw() + 
  draw_label(
    "Some classes are very large (class size means marked)",
    x = 0,
    hjust = 0
  ) +
  theme(
    plot.margin = margin(0, 0, 0, 7)
  )
plot_grid(title,
          combined,
          legend,
          ncol = 1,
          rel_heights = c(.1, 2))
```

@@

Interestingly, there is a large difference between how students and instructors rate perceived similarity and ISR. Students, on average, rate end-of-term similarity at 3.6, while instructors give an average rating of 2.7. Students give an average end-of-term ISR rating of 4.1, while instructors give an average of only 3.5. This could be due to more positive student perceptions, or due to students giving higher ratings for equivalent perceptions. Both groups rate ISR more highly than similarity, but it is difficult to compare the weight of these scores as the scales are not necessarily measuring in the same units.

```{r continuousoutcomes, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6, fig.height= 10, fig.cap="Histograms illustrating continous and ordinal key outcomes"}
# Plot initial student similarity perception
s1_simplot <- expdata %>% 
  ggplot(aes(x = s1_sim, fill = treatment)) +
  geom_histogram(position = "dodge") +
  labs(x = "Initial student similarity perception", y = "Frequency", fill = "Group") +
  scale_x_continuous(limits = c(.8,5.2)) +
  scale_fill_manual(values = c("#52f2b9", "#868cf7")) +
  theme_minimal()

# Plot initial student ISR perception
s1_tsrplot <- expdata %>% 
  ggplot(aes(x = s1_tsr, fill = treatment)) +
  geom_histogram(position = "dodge") +
  labs(x = "Student ISR anticipation", y = "Frequency") +
  scale_x_continuous(limits = c(.8,5.2)) +
  scale_fill_manual(values = c("#52f2b9", "#868cf7")) +
  theme_minimal()

# Plot end-of-term student similarity perception
s2_simplot <- expdata %>% 
  ggplot(aes(x = s2_sim, fill = treatment)) +
  geom_histogram(position = "dodge") +
  labs(x = "End-of-term student similarity perception", y = "Frequency") +
  scale_x_continuous(limits = c(.8,5.2)) +
  scale_fill_manual(values = c("#52f2b9", "#868cf7")) +
  theme_minimal()

# PLot end of term student ISR perception
s2_tsrplot <- expdata %>% 
  ggplot(aes(x = s2_tsr, fill = treatment)) +
  geom_histogram(position = "dodge") +
  labs(x = "End-of-term student ISR perception", y = "Frequency") +
  scale_x_continuous(limits = c(.8,5.2)) +
  scale_fill_manual(values = c("#52f2b9", "#868cf7")) +
  theme_minimal()

# Plot end of term instructor similarity perception
t2_sim1plot <- expdata %>% 
  ggplot(aes(x = t2_sim1, fill = treatment)) +
  geom_histogram(position = "dodge") +
  labs(x = "End-of-term instructor similarity perception", y = "Frequency") +
  scale_x_continuous(limits = c(.8,5.2)) +
  scale_fill_manual(values = c("#52f2b9", "#868cf7")) +
  theme_minimal()

# Plot end of term instructor ISR perception
t2_tsrplot <- expdata %>% 
  ggplot(aes(x = t2_tsr, fill = treatment)) +
  geom_histogram(position = "dodge") +
  labs(x = "End-of-term instructor ISR perception", y = "Frequency") +
  scale_x_continuous(limits = c(.8,5.2)) +
  scale_fill_manual(values = c("#52f2b9", "#868cf7")) +
  theme_minimal()

# Plot grade
gradeplot <- expdata %>% 
  ggplot(aes(x = grade, fill = treatment)) +
  geom_histogram(position = "dodge") +
  labs(x = "Course grade", y = "Frequency") +
  scale_x_continuous(limits = c(-.1,4.5)) +
  scale_fill_manual(values = c("#52f2b9", "#868cf7")) +
  theme_minimal()

# plot standardized grades
std_gradeplot <- expdata %>% 
  ggplot(aes(x = std_grade, fill = treatment)) +
  geom_histogram(position = "dodge") +
  labs(x = "Standardized course grade", y = "Frequency") +
  scale_x_continuous(limits = c(-4.5,3)) +
  scale_fill_manual(values = c("#52f2b9", "#868cf7")) +
  theme_minimal()

# Plot final exam grades
t2_finalexamplot <- expdata %>% 
  ggplot(aes(x = t2_finalexam, fill = treatment)) +
  geom_histogram(position = "dodge") +
  labs(x = "Final grade", y = "Frequency") +
  scale_x_continuous(limits = c(-.1,4.5)) +
  scale_fill_manual(values = c("#52f2b9", "#868cf7")) +
  theme_minimal()

# Combine plots in a grid
combined <- plot_grid(s1_simplot + theme(legend.position = "none"), s1_tsrplot + theme(legend.position = "none"),
          s2_simplot + theme(legend.position = "none"), s2_tsrplot + theme(legend.position = "none"),
          t2_sim1plot + theme(legend.position = "none"), t2_tsrplot + theme(legend.position = "none"),
          std_gradeplot + theme(legend.position = "none"), t2_finalexamplot + theme(legend.position = "none"),
          ncol = 2)
legend <- get_legend(s1_simplot + 
                       guides(color = guide_legend(nrow = 1)) +
                       theme(legend.position = "bottom"))
title <- ggdraw() + 
  draw_label(
    "Most key outcomes unaffected by treatment",
    x = 0,
    hjust = 0
  ) +
  theme(
    plot.margin = margin(0, 0, 0, 7)
  )
plot_grid(title,
          combined,
          legend,
          ncol = 1,
          rel_heights = c(.1, 2, .1))
```

@@

From the 36,838 observations, or potential units of study (corresponding to all undergraduate student records of the university), only 2,273 were used in analysis. Units of study were excluded because the instructor did not participate in the study (33,483); the instructor or course wasn't actually eligible (96); the instructor administered the wrong survey (31); the student did not consent  (116); the student did not complete the first survey (304); the instructor did not complete the second survey (107); or one participant did not spend more than a second reading the consent page and more than ten seconds reading each page with five questions or more (428). The time limits are somewhat arbitrary, and different choices could have different analysis results.

@@

$X_{1i}$ is a vector of student-level covariates (student ISR anticipation, gender, CGPA, year of study). Different covariates are used to model different outcomes. The original study used gender and CGPA as controls for grade based outcomes only. @citeoriginal state this is because females earn higher grades generally. This paper uses them as controls for the majority of outcomes. This is because of the potential link between student performance and perceptions of similarity and ISR. ISR anticipation is used as a control only for end-of-term student ISR perception.

$X_{2j}$ is a vector of instructor-level covariates (class size). The original study considered instructor fixed effects instead. Because each participating instructor taught one class in the study, the class sizes correspond directly to the instructor IDs, so it was inadvisable to use both. This paper selected class size because of the larger potential for transferable insights. These particular instructors are not of general interest, but the effect of class size may be.

$\epsilon_{ij}$ is a standard error. As this paper does not use clustered standard errors like @citeoriginal, there are slight differences in standard errors and p-values when compared with the original paper. This does not affect the regression coefficients.

@@

Equation \@ref(eq:linmodel) is @citeoriginal's linear model, used for continuous outcomes. This includes complete scale outcomes: initial student similarity perception, end-of-term student similarity perception, end-of-term student ISR perception, and end-of-term instructor ISR perception. Because these scales are created using 6-7 questions answered on a scale of 1 to 5, they have a total of 25 or 30 possible values, and they can be treated as continuous variables and modeled using linear functions. It also includes grade-based outcomes: course grade and objectively graded exam grade, both of which are shown on a 4.0 GPA scale.

@@

Equation \@ref(eq:ordlogmodel) is @citeoriginal's ordinal logistic model, used for the ordinal outcome, which is end-of-term instructor similarity perception.The instructors were only asked to answer one five-point scale question about similarity, yielding an ordinal scale rather than a pseudo-continuous scale. Therefore, this outcome is treated differently and the model selected is specifically for ordinal outcomes, where the variable is ordered but not equally spaced. Because this variable is measured differently than the two student similarity perception scales, and therefore modeled differently, direct comparisons between student and instructor perception of similarity should be made with caution.

@@

When creating additional models, I focused on three key outcomes: standardized course grade, end-of-term student perceived ISR scale, and end-of-term student perceived similarity scale. I selected these because interventions like @citeoriginal's are aimed at improving student outcomes (like course performance), and it is much more likely to detect an impact in the immediate course outcomes as opposed to outcomes in subsequent terms; and because previous studies focus more on student perception than instructor perception. Equation \@ref(eq:mylinmodel) shows the linear model used. $X_{3jk}$ represents student-instructor commonality covariates (matching racial self-ID, matching gender self-ID, age difference).

@@

In Table \@ref(tab:modeltable), Model 1 shows that the treatment had a statistically significant effect on initial student similarity perception ($\beta_{1}$ = 0.194, 95% CI [0.136, 0.253]). Similarly, Model 2 shows that the treatment had a statistically significant effect on  student end-of-term similarity perception ($\beta_{1}$ = 0.11, 95% CI [0.038, 0.183]), though this effect size is smaller. Both coefficients are minimal on a scale of 1 to 5. The adjusted R-squared values, which indicate the proportion of variation in the outcome that can be explained by the treatment and covariates, are low (0.032 and 0.022). These models are helpful for analyzing treatment effect, but they do not aim to be and are not useful predictors of student similarity perception, not even within the sample.

@@

Models 3 and 4 show that student ISR perception is significantly associated with course grade, while instructor ISR perception is significantly associated with course grade and final grade. Instructor perception is more strongly associated ($\Gamma$ = 0.34, 95% CI [0.295, 0.385] for course grade; $\Gamma$ = 0.643, 95% CI [0.642, 0.759] for final grade) than student perception ($\Gamma$ = 0.115, 95% CI [0.064, 0.164] for course grade).

@@

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# create variables to indicate matching student-instructor traits
expdata <- expdata %>%
  mutate(race_match = ifelse(t1_race_1 == s1_race_1
                             & t1_race_2 == s1_race_2 
                             & t1_race_3 == s1_race_3 
                             & t1_race_4 == s1_race_4 
                             & t1_race_5 == s1_race_5 
                             & t1_race_6 == s1_race_6 
                             & t1_race_7 == s1_race_7,
                             1,
                             0)) %>%
  mutate(gend_match = ifelse(t_female == s1_female,
                             1,
                             0)) %>%
  mutate(age_dif = t_age - s1_age)
```
